---
title: "Take-home Exercise 1"
description: |
  In this Take-home Exercise, I will explore using appropriate static statistical graphic methods to understand the demographic of the city of Engagement, Ohio USA.
author:
  - name: Jeremiah Lam 
    url: https://sg.linkedin.com/in/jeremiah-lam-6156238a
    affiliation: School of Computing and Information Systems
    affiliation_url: https://scis.smu.edu.sg/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this take-home exercise, appropriate static statistical graphic methods are used to reveal the demographic of the city of Engagement, Ohio USA. 

The data is processed by using appropriate tidyverse family of packages and the statistical graphics are prepared using ggplot2 and its extensions.

# Getting Started

Before getting started, it is important to ensure that the required R packages have been installed. If yes, we will load the R packages. If they have yet to be installed, we will install the R packages and load them onto R environment. 

The chunk code below will do the trick.

```{r}
packages = c('tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# Importing Data

The code chunk below imports *Participants.csv* from the data folder into R by using [`read_csv()`](https://readr.tidyverse.org/reference/read_delim.html) of [**readr**](https://readr.tidyverse.org/reference/index.html) package and saves it as a tibble data frame called *Participants*.

```{r}
Participants <- read_csv("data/Participants.csv")
```

# A Simple Bar Chart

```{r}
ggplot(data = Participants,
       aes(x = age)) +
  geom_bar()
```

